### 2023-05-20 episodes/data-visualization.Rmd

---
title: 'Data Visualization'
teaching: 10
exercises: 60
---

## Comparison of clustering results

::::::::::::::::::::::::::::::::::::::: objectives

- Explore the resulting trees in conjunction with the meta data.
- Make an estimation on the likelihood of transmission events

::::::::::::::::::::::::::::::::::::::::::::::::::

:::::::::::::::::::::::::::::::::::::::: questions

- How are phylogenetic trees viewed and compared?
- How can I visualize several layers of data?

::::::::::::::::::::::::::::::::::::::::::::::::::

We grouped (clustered) our isolates by the information that we extracted from the sequencing reads. We compared our isolates to a reference genome and clustered them on basis of the single nucleotide variations that they exhibit when compared to this reference.

![](fig/Workflow.png){alt='workflow'}

Next, we will visualize this clustering in the context of the meta data. Make sure you have the SNP tree that you generated (or can be downloaded here [core.aln.newick](files/core.aln.newick)) on your own computer. Download also the [metadata file](files/Molepi_meta.csv).

## Visualization of genetic information and metadata

Visualization is frequently used to aid the interpretation of complex datasets. Within microbial genomics, visualizing the relationships between multiple genomes as a tree provides a framework onto which associated data (geographical, temporal, phenotypic and epidemiological) are added to generate hypotheses and to explore the dynamics of the system under investigation.

1. Open Chrome. Go to [microreact](https://microreact.org/). 
2. Click on 'Upload'. Drag-and-drop the newick tree produced by snippy and IQ-TREE ([core.aln.newick](files/core.aln.newick)). 
3. Click on "Add more files" and select the the [metadata file](files/Molepi_meta.csv), which is a table comprising metadata file. Choose 'Data (csv or tsv)' as file kind.
4. Click on continue.

First, the tree should be rooted with the reference strain (the one that has the longest branch). Hover over the middle of the long branch, until a circle appears. Right-click on it, and select "Set as root (re-root)".

Explore the location, time and further meta data. Play around with the different visualization options with the help of the [documentation of microreact](https://docs.microreact.org/).

::::::::::::::::::::::::::::::::::::::  discussion

## Discussion

1. Which transmission events are likely based on the metadata (location, RFLP, date) alone?
2. Which transmission events are likely based on the SNP data?
3. Draw a transmission tree if possible.

::::::::::::::::::::::::::::::::::::::::::::::::::

:::::::::::::::::::::::::::::::::::::::: keypoints

- Genetic information can confirm or contradict the meta data

::::::::::::::::::::::::::::::::::::::::::::::::::