### 2023-05-22 episodes/msa.Rmd

---
title: 'Multiple sequence alignment'
teaching: 10
exercises: 30
---

:::::::::::::::::::::::::::::::::::::: questions

- How do you align multiple sequences?
- Why is it important to properly align sequences?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Use `mafft` to align multiple sequences.
- Test different algorithms.

::::::::::::::::::::::::::::::::::::::::::::::::

## Introduction

::::::::::::::::::::::::::::::::::::: challenge

## Challenge 1: Can you do it?

What is the output of this command?

```r
paste("This", "new", "lesson", "looks", "good")
```

:::::::::::::::::::::::: solution

## Output

```output
[1] "This new lesson looks good"
```

:::::::::::::::::::::::::::::::::

## Challenge 2: how do you nest solutions within challenge blocks?

:::::::::::::::::::::::: solution

You can add a line with at least three colons and a `solution` tag.

:::::::::::::::::::::::::::::::::
::::::::::::::::::::::::::::::::::::::::::::::::

## Figures

You can include figures generated from R Markdown:

```{r pyramid, fig.alt = "pie chart illusion of a pyramid", fig.cap = "Sun arise each and every morning"}
pie(
  c(Sky = 78, "Sunny side of pyramid" = 17, "Shady side of pyramid" = 5), 
  init.angle = 315, 
  col = c("deepskyblue", "yellow", "yellow3"), 
  border = FALSE
)
```
Or you can use pandoc markdown for static figures with the following syntax:

`![optional caption that appears below the figure](figure url){alt='alt text for
accessibility purposes'}`

![You belong in The Carpentries!](https://raw.githubusercontent.com/carpentries/logo/master/Badge_Carpentries.svg){alt='Blue Carpentries hex person logo with no text.'}

## Math

One of our episodes contains $\LaTeX$ equations when describing how to create
dynamic reports with {knitr}, so we now use mathjax to describe this:

`$\alpha = \dfrac{1}{(1 - \beta)^2}$` becomes: $\alpha = \dfrac{1}{(1 - \beta)^2}$

Cool, right?

::::::::::::::::::::::::::::::::::::: keypoints

- Use `.md` files for episodes when you want static content
- Use `.Rmd` files for episodes when you need to generate output
- Run `sandpaper::check_lesson()` to identify any issues with your lesson
- Run `sandpaper::build_lesson()` to preview your lesson locally

::::::::::::::::::::::::::::::::::::::::::::::::