### 2023-05-22 episodes/introduction.Rmd
First part of MSA ready

---
title: 'Introduction'
teaching: 20
exercises: 10
---

:::::::::::::::::::::::::::::::::::::: questions

- How does evolution work (in microbes)?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Test your prior knowledge about evolutionary biology.
- Discuss a few "difficult" questions about evolution, with a specific focus on 
how microbes evolve.

::::::::::::::::::::::::::::::::::::::::::::::::

## Prior knowledge

This is an ice-breaking exercise. Don't panic.

:::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::: instructor

Let the students discuss two and two for 4-5 minutes. Encourage them to write
keywords on the board.

::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: challenge

## What does evolution need to work

- Discuss with your neighbor for 5 minutes
- Write up to three keywords on the board

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: challenge

## Quiz about evolution

Answer the quiz as you follow along, [using this link](https://pollev.com/lionelguy323/) or by using the following QR code ![QR code](fig/qr_code_quiz.png){alt='QR code linking to the interactive quiz'}

:::::::::::::::::::::::::::::::::::::: solution

The answers will be discussed in the presentation later.

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: keypoints

- Learn the very basic principle of evolution.
- Evolution can be counter-intuitive.

::::::::::::::::::::::::::::::::::::::::::::::::