### 2024-03-13 code/templates.Rmd
Shoved some template text to a Rmd file

::::::::::::::::::::::::::::::::::::: challenge

```r
paste("This", "new", "lesson", "looks", "good")
```
:::::::::::::::::::::::: solution

## Output

```output
[1] "This new lesson looks good"
```

:::::::::::::::::::::::::::::::::

## Challenge 2: how do you nest solutions within challenge blocks?

:::::::::::::::::::::::: solution

You can add a line with at least three colons and a `solution` tag.

:::::::::::::::::::::::::::::::::
::::::::::::::::::::::::::::::::::::::::::::::::