### 2024-03-26 episodes/msa.Rmd
Finalizing Blast, upgrading MSA

---
title: 'Multiple sequence alignment'
teaching: 0
exercises: 240
---

:::::::::::::::::::::::::::::::::::::: questions

- How do you align multiple sequences?
- Why is it important to properly align sequences?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Use `mafft` to align multiple sequences.
- Test different algorithms.

::::::::::::::::::::::::::::::::::::::::::::::::

## Introduction

In this episode, we are exploring multiple sequence alignment (MSA). We are going to use [`mafft`](https://mafft.cbrc.jp/alignment/software/algorithms/algorithms.html) to align homologs of [RpoB](https://en.wikipedia.org/wiki/RpoB), the β subunit of the bacterial RNA polymerase. It is a long, multi-domain protein, suitable for showing issues related to MSA.

You will also

But first, go to your own folder and create a phylogenetics subfolder

## Task 1: Use different flavors of `MAFFT` and compare the results

Start by finding the data, which is a [fasta file](episodes/files/rpoB.fasta) containing 19 sequences of RpoB, gathered by aligning RpoB from *E. coli* to a very small database present at NCBI, the [landmark](https://blast.ncbi.nlm.nih.gov/smartblast/smartBlast.cgi?CMD=Web&PAGE_TYPE=BlastDocs) database. These sequences are a subset of the sequences gathered in the previous exercise on `BLAST`.

The file is available in `/proj/g2020004/nobackup/3MK013/data/`.

:::::::::::::::::::::::::::::::::::::::  challenge

## Challenge 1.1: prepare the terrain

Go to your own folder, use the `mkdir` command to make a `phylogenetics` subfolder, and move into it. Also, start the `interactive` session.

:::::::::::::::  solution

Remember to replace `<username>` by your name.

```bash
$ interactive -A uppmax2024-2-10 -M snowy -t 4:00:00
$ cd /proj/g2020004/nobackup/3MK013/<username>
$ mkdir phylogenetics
$ cd phylogenetics
```

:::::::::::::::::::::::::

## Challenge 1.2: copy the file

Copy the file to your newly created `~/phylogenetics/` folder.

:::::::::::::::  solution

```bash
$ cp ../../data/rpoB.fasta .
```

:::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::::

### Renaming sequences

Look at the accession ids of the fasta sequences: they are not very informative:

```bash
$ grep '>' rpoB.fasta | head -n 5
```

```output
>WP_000263098.1 MULTISPECIES: DNA-directed RNA polymerase subunit beta [Enterobacteriaceae]
>WP_011070610.1 DNA-directed RNA polymerase subunit beta [Shewanella oneidensis]
>WP_003114335.1 DNA-directed RNA polymerase subunit beta [Pseudomonas aeruginosa]
>WP_002228870.1 DNA-directed RNA polymerase subunit beta [Neisseria meningitidis]
>WP_003436174.1 MULTISPECIES: DNA-directed RNA polymerase subunit beta [Eubacteriales]
```

Keep the taxonomic name following the accession id, separated by `_`, using a bit of `sed` magic. Save the resulting file for later use, and show the headers again.

```bash
$ sed "/^>/ s/>\([^ ]*\) \([^\[]*\)\[\(.*\)\]/>\\3_\\1/"  rpoB.fasta | sed "s/ /_/g" > rpoB_renamed.fasta
$ grep '>' rpoB_renamed.fasta | head -n 5
```

::: discussion

## Nerd alert: dissecting the `sed` magic

This is optional reading.

The [`sed`](https://www.gnu.org/software/sed/manual/sed.html) command matches (and possibly substitutes) strings (chains of characters). In that case, the goal is to simplify the header by putting the taxonomic group first but keeping it informative enough by adding the accession number. The strategy is to match what is between the `>` and the first space, then what is between square parentheses, and put them back, separated by an underscore.

The `sed` command first matches only lines that start with a `>` (`/^>/`). It then substitutes (general pattern `s/<something>/<something else>/`) a text with another one. The first part is to match the accession id, between (escaped) brackets, which comes after the `>` at the beginning of the line. This is expressed as `>\([^ ]*\) `: match any number of non-space characters and put it in memory. Then, the description is matched by `\([^\[]*\)`, any number of characters that are not an opening bracket `[`, and put into memory. Finally, the taxonomic description is matched: `\[\(.*\)\]`, that is, any number of characters between square brackets is stored into memory. The whole line is then replaced with a `>`, the third match into memory, followed by an `_` and the content of the first match into memory `>\\3_\\1`. Then, all the input is passed through sed again, to replace any space with an underscore: `s/ /_/g` and the output is stored in a different file.

::::::::::::::

```output
>Enterobacteriaceae_WP_000263098.1
>Shewanella_oneidensis_WP_011070610.1
>Pseudomonas_aeruginosa_WP_003114335.1
>Neisseria_meningitidis_WP_002228870.1
>Eubacteriales_WP_003436174.1
```

It looks better.

### Alignment with progressive algorithm

Use `mafft` with a progressive algorithm to align the sequences.

:::::::::::::::::::::::::::::::::::::::  challenge

## Challenge 1.3: Run `mafft` with progressive algorithm

Use the FFT-NS-2 algorithm from `mafft` to align the renamed sequences. Also, record the time it takes for `mafft` to complete the task.

Use the `module` command to load `bioinfo-tools` and `MAFFT`.

:::::::::::::::  solution

```bash
$ module load bioinfo-tools MAFFT
$ time mafft-fftns rpoB_renamed.fasta > rpoB.fftns.aln
```

```output
...
Strategy:
 FFT-NS-2 (Fast but rough)
 Progressive method (guide trees were built 2 times.)

If unsure which option to use, try 'mafft --auto input > output'.
For more information, see 'mafft --help', 'mafft --man' and the mafft page.

The default gap scoring scheme has been changed in version 7.110 (2013 Oct).
It tends to insert more gaps into gap-rich regions than previous versions.
To disable this change, add the --leavegappyregion option.

mafft-fftns rpoB_renamed.fasta > rpoB.fftns.aln  1.02s user 0.52s system 104% cpu 1.471 total
```

The last line is the output of the `time` command. It took 1.47 seconds to complete this time.

:::::::::::::::::::::::::

::: hint

Type `mafft` and try tab-complete to see all versions of `mafft`.

Try the command `time`

:::

::::::::::::::::::::::::::::::::::::::::::::::::::

### Alignment with iterative algorithm

Now use one of the supposedly better iterative algorithm of `mafft` to align the same sequences. Choose the [E-INS-i algorithm](https://mafft.cbrc.jp/alignment/software/algorithms/algorithms.html) which is suited for proteins that have highly conserved motifs interspersed with less conserved ones.

:::::::::::::::::::::::::::::::::::::::  challenge

## Challenge 1.4

Use the E-INS-i algorithm from mafft to align the renamed sequences. Also, record the time it takes for `mafft` to complete the task.

:::::::::::::::  solution

```bash
$ time mafft-einsi rpoB_renamed.fasta > rpoB.einsi.aln
```

```output
...
Strategy:
 E-INS-i (Suitable for sequences with long unalignable regions, very slow)
 Iterative refinement method (<16) with LOCAL pairwise alignment with generalized affine gap costs (Altschul 1998)

If unsure which option to use, try 'mafft --auto input > output'.
For more information, see 'mafft --help', 'mafft --man' and the mafft page.

The default gap scoring scheme has been changed in version 7.110 (2013 Oct).
It tends to insert more gaps into gap-rich regions than previous versions.
To disable this change, add the --leavegappyregion option.

Parameters for the E-INS-i option have been changed in version 7.243 (2015 Jun).
To switch to the old parameters, use --oldgenafpair, instead of --genafpair.

mafft-einsi rpoB_renamed.fasta > rpoB.einsi.aln  6.84s user 0.65s system 99% cpu 7.542 total
```

It now took 7.54 seconds to complete this time, i.e. about 5 times slower than with the progressive algorithm. It doesn't make a big difference now, but with hundreds of sequences it will make one.

:::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::::

### Alignment visualization

You will now inspect the two resulting alignment methods. There are no convenient way to do that from Uppmax, and the easiest solution is to download the alignments **on your computer** and to use either [`seaview`](https://doua.prabi.fr/software/seaview) (you will need to install it on your computer) or an online alignment viewer like [AlignmentViewer](https://alignmentviewer.org/).

Arrange the two windows on top of each other. Change the fontsize (Props -> Fontsize in Seaview) to 8 to see a larger portion of the alignment.

:::::::::::::::::::::::::::::::::::::::  challenge

Can you spot differences? Which alignment is longer?

Hint: try to scroll to position 800-900. What do you see there? How are the blocks arranged?

:::::::::::::::  solution

![](episodes/fig/seaview.png){alt='Alignments shown in seaview'}

::::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::::

If you have time over, spend it exploring the different options of Seaview/AlignmentViewer.

## Task 2: Trim the alignment

Often, some regions of the alignment don't align properly, either because they contain low complexity segments (hinges in proteins) or evolved through repeated insertions/deletions, which alignment program cannot handle properly. It is thus good practice to remove (trim) these regions, as they are likely to worsen the quality of the subsequent phylogenetic trees. On the other hand, trimming too much of the alignment removes also potentially valuable information. There is thus a balance to be found.

In this part, you will use two different alignemnt trimmers, (`TrimAl`)[http://trimal.cgenomics.org/] and (`ClipKIT`)[https://jlsteenwyk.com/ClipKIT/].

### Trimming with TrimAl

First, look at the list of options available with `trimAl`