### 2024-03-27 episodes/phylogenetics.Rmd
Switched the phylogenetics episode to Uppmax and phylo.io

---
title: 'Phylogenetics'
teaching: 0
exercises: 240
---

:::::::::::::::::::::::::::::::::::::: questions

- How do you build a simple, distance-based phylogenetic tree?
- How do you build a phylogenetic tree with more advanced methods?
- How do you ascertain statistical support for phylogenetic trees?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Learn the basic of phylogenetics tree building, taking the simplest of the examples with a UPGMA tree.
- Learn how to use phylogenetics algorithms, neighbor-joining and maximum-likelihood.
- Learn how to perform and show bootstraps.

::::::::::::::::::::::::::::::::::::::::::::::::

## Exercise 1: Paper-and-pen phylogenetic tree

### Setup

The exercise is done for a large part with pen and paper, and then a demonstration in R on your laptop. You can already RStudio. We'll also use the R package [`ape`](https://emmanuelparadis.github.io/) , which you should install if it's not present on your setup:

```r
install.packages('ape')
```

And to load it:

```{r}
library(ape)
```

### UPGMA-based tree

Load the tree in fasta format, reading from a `temp` file
```{r}
FASTAfile <- tempfile("aln", fileext = ".fasta")
cat(">A", "----ATCCGCTGATCGGCTG----",
    ">B", "GCTGATCCGTTGATCGG-------",
    ">C", "----ATCTGCTCATCGGCT-----",
    ">D", "----ATTCGCTGAACTGCTGGCTG",
    file = FASTAfile, sep = "\n")
aln <- read.dna(FASTAfile, format = "fasta")
```
Now look at the alignment. Notice there are gaps, which we don't want in this example. We also remove the non-informative (identical) columns.

```{r}
alview(aln)
aln_filtered <- aln[,c(7,8,10,12,14,16)]
alview(aln_filtered)
```
Now we have a simple alignment as in the lecture. Dots (`.`) mean that the sequence is identical to the top row, which makes it easier to calculate distances.

::::::::::::::::::::::::::::::::::::: challenge

### Calculating distance "by hand"

Let's use a matrix to calculate distances between sequences. We'll just count the number of differences between the sequences. We'll then group the two closest sequences. Which are they?

Table: Distances between the sequences.

|   | A  | B  | C  | D  |
| - | -: | -: | -: | -: |
| A |    |    |    |    |
| B |    |    |    |    |
| C |    |    |    |    |
| D |    |    |    |    |

:::::::::::::::::::::::: solution

Here is the solution:

Table: Distances between the sequences.

|   | A  | B  | C  | D  |
| - | -: | -: | -: | -: |
| A |    |    |    |    |
| B |  1 |    |    |    |
| C |  2 | 3  |    |    |
| D |  3 | 4  | 5  |    |

The two closest sequences are A and B.

:::::::::::::::::::::::::::::::::

Let's now cluster together A and B, and calculate the average distance from AB to the other sequences, weighted by the size of the clusters.

Table: Recalculated distances.

|    | AB  | C  | D  |
| -  | -:  | -: | -: |
| AB |     |    |    |
| C  |     |    |    |
| D  |     |    |    |

:::::::::::::::::::::::: solution

The average distance for AB to C is calculated as follow:

$d(AB,C) = \dfrac{d(A,C) + d(B,C)}{2} = \dfrac{2 + 3}{2} = 2.5$

And so on for the other distances:

Table: Recalculated distances.

|    | AB  | C  | D  |
| -  | -:  | -: | -: |
| AB |     |    |    |
| C  | 2.5 |    |    |
| D  | 3.5 |  5 |    |

:::::::::::::::::::::::::::::::::

Now the shortest distance is AB to C. Let's recalculate the distance to D again.

Table: Recalculated distances.

|    | ABC | D  |
| -  | -:  | -: |
| ABC|     |    |
| D  |     |    |

:::::::::::::::::::::::: solution

$d(ABC,D) = \dfrac{d(AB,D) * 2 + d(C,D) * 1}{2 + 1} = \dfrac{3.5 * 2 + 5 * 1}{3} = 4$

Table: Recalculated distances.

|    | ABC | D  |
| -  | -:  | -: |
| ABC|     |    |
| D  | 4   |    |

:::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::

The tree is reconstructed by dividing the distances equally between the two leaves. 
- A-B: each 0.5.
- AB-C: each side gets 2.5/2 = 1.25. The branch to AB is 1.25 - 0.75 = 0.75
- ABC-D: each side gets 4/2 = 2. The branch to ABC is 2 - 0.75 - 0.5 = 0.75

![UPGMA tree](fig/upgma_manual.png){alt='Manually built UPGMA tree'}

Let's know do the same using bioinformatics tools.

::::::::::::::::::::::::::::::::::::: challenge

We'll use `dist.dna` to 
calculate the distances. We'll use a "N" model, that just counts the differences and doesn't correct or normalizes. We'll use the function `hclust` to perform the UPGMA method calculation. The tree is then plotted, and the branch lengths plotted with `edgelabels`:

:::::::::::::::::::::::: solution

```{r}
dist_matrix <- dist.dna(aln_filtered, model="N")
dist_matrix
tree <- as.phylo(hclust(dist_matrix, "average"))
plot(tree)
edgelabels(tree$edge.length)

```

:::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::

## Exercise 2: maximum-likelihood tree

### Introduction

In the previous episode, we inferred a simple phylogenetic tree using UPGMA, without correcting the distance matrix for multiple substitutions. UPGMA has many shortcomings, and gets worse as distance increases. Here we'll test Neighbor-Joining, which is also a distance-based, and a maximum-likelihood based approach with [IQ-TREE](http://www.iqtree.org/).

### Neighbor-joining

The principle of [Neighbor-joining method](https://en.wikipedia.org/wiki/Neighbor_joining) (NJ) is to start from a star-like tree, find the two branches that, if joined, minimize the total tree length, join then, and repeat with the joined branches and the rest of the branches.

::::::::::::::::::::::::::::::::::::: challenge

To perform NJ on our sequences, we'll use the function inbuilt in Seaview. First (re-)open the alignment we obtained from `mafft` with the E-INS-i method, trimmed with ClipKIT.

:::::::::::::::::::::::: solution

```bash
seaview rpoB.einsi.clipkit.aln &
```

:::::::::::::::::::::::::::::::::

In the Seaview window, select Trees -> Distance methods. Keep the BioNJ option ticked. BioNJ is an updated version of NJ, more accurate for longer distances. Tick the "Bootstrap" option and leave it at 100. We'll discuss these later. Then click on "Go".

::::::::::::::::::::::::::::::::::::::::::::::::

What do you see? Is the tree rooted? Is the pattern species coherent with what you know of the tree of life? Any weird results?

::::::::::::::::::::::::::::::::::::: challenge

Redo the BioNJ tree for the other alignment we inferred.

:::::::::::::::::::::::: solution

```bash
seaview rpoB.fftns.aln &
```

:::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::

Do you see any differences between the two trees? What can you make out of it?

### Maximum likelihood

We will now use IQ-TREE to infer a maximum-likelihood (ML) tree of the RpoB dataset we aligned with `mafft` previously. While you performed the first trees on your laptop, you will infer the ML tree on Uppmax. As usual, use the `interactive` command to gain access to compute time. Go to the `phylogentics` folder created in the MSA episode, and load the right modules.

:::::: challenge

Get to the right folder, require compute time and load the right modules.

::: solution

 ```bash
interactive -A uppmax2024-2-10 -M snowy -t 4:00:00
module load bioinfo-tools iqtree
cd /proj/g2020004/nobackup/3MK013/<username>/phylogenetics
```

::::::::::::

::::::::::::::::

Loading the `iqtree` module resulted in:

```output
iqtree/2.2.2.6-omp-mpi: executables are prefixed with 'iqtree2'. Additional modules are required for the multi-node MPI version. See 'module help iqtree/2.2.2.6-omp-mpi'.
```

You won't be using the multi-node MPI version, so you don't need to load additional modules. But you will need to use `iqtree2` instead of the "normal" `iqtree`.

Then run your first tree, on the `ClipKIT`-trimmed alignment.

```bash
$ iqtree2 -s rpoB.einsi.clipkit.aln -m TEST -B 1000
```

Here, we tell IQ-TREE to use the alignment `rpoB.einsi.clipkit.aln`, and to test among the standard substitution models which one fits best (`-m TEST`). We also tell IQ-TREE to perform 1000 ultra-fast bootstraps (`-B 1000`). We'll discuss these later.

IQ-TREE is a very complete program that can do a large variety of phylogenetic analysis. To get a flavor of what it's capable of, look at its [extensive documentation](http://www.iqtree.org/doc/).

Have a look at the output files:

```bash
ls rpoB.einsi.clipkit.aln.*
```

```output
rpoB.einsi.clipkit.aln.bionj    rpoB.einsi.clipkit.aln.iqtree  rpoB.einsi.clipkit.aln.model.gz
rpoB.einsi.clipkit.aln.ckp.gz   rpoB.einsi.clipkit.aln.log     rpoB.einsi.clipkit.aln.splits.nex
rpoB.einsi.clipkit.aln.contree	rpoB.einsi.clipkit.aln.mldist  rpoB.einsi.clipkit.aln.treefile
```

There are two important files:  
* `*.iqtree` file provides a text summary of the analysis. 
* `*.treefile` is the resulting tree in [Newick format](https://en.wikipedia.org/wiki/Newick_format).

To visualize the tree, open the [Beta version of phylo.io](https://beta.phylo.io/viewer/#). Click on "add a tree" and copy/paste the content of the `treefile` (use e.g. `less` or `cat` to display it) in the box. Make sure the "Newick format" is selected and click on "Done".

The tree appears as unrooted. It is good practice to start by ordering the nodes and root it. There is no good way to automatically order the branches in phylo.io as of yet, but rerooting can be done by clicking on a branch and selecting "Reroot". Reroot the tree between archaea and bacteria. Now the tree makes a bit more sense.

Scrutinize the tree. Is it different from the BioNJ tree generated in Seaview? How?

::::::::::::::::::::::::::::::::::::: challenge

Redo a ML tree from the other alignment (FFT-NS) we inferred with `mafft` and display the resulting tree in FigTree

:::::::::::::::::::::::: solution

```bash
$ iqtree2 -s rpoB.fftns.aln -m TEST -B 1000
$ figtree rpoB.fftns.aln.treefile &
```

:::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::

### Bootstraps

We have inferred four trees:
* Two based on the alignment generated from the E-INS-i algorithm (trimmed by ClipKIT), two from the FFT-NS.
* Two inferred with the BioNJ algorithm and two with the ML algorithm (IQ-Tree)

Along the way, we've generated bootstraps for all our trees. Now show them on all four trees.
* For the Seaview trees, tick the 'Br support' box
* For the trees shown in phylo.io, click on "Settings" > "Branch & Labels" and above the branch, click the drop-down menu and select "Data".

::::::::::::::::::::::::::::::::::::: challenge

Compare all four trees. Do you find any significant differences?

Hint: what are *Glycine* and *Arabidopsis*? What about *Synechocystis* and *Microcystis*?

Hint: search for the accession number of the *Glycine* RpoB sequence on [NCBI](https://www.ncbi.nlm.nih.gov/search/). Any hint there?

:::::::::::::::::::::::: solution

The two plant RpoB sequences are from chloroplastic genomes, and thus branch together with the cyanobacterial sequences, in some of the phylogenies at least.

:::::::::::::::::::::::::::::::::

What about the support values for grouping these two groups? How high are they?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: keypoints

- The simplest of the trees are distance-based.
- UPGMA works by clustering the two closest leaves and recalculating the distance matrix.
- Neighbor-joining is distance-based and fast, but not necessarily very accurate
- Maximum-likelihood is slower, but more accurate

::::::::::::::::::::::::::::::::::::::::::::::::

## Exercise 3: Genetic drift

As a practical way to understand genetic drift, let's play with population size, selection coefficients, number of generations, etc.

:::::::::::::::::::::::::::::::::::::: questions

- How does random genetic drift influence fixation of alleles?
- What is the influence of population size?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Understand how population size influences the probability of fixation of an allele.
- Understand how slightly deleterious mutations may get fixed in small populations.

::::::::::::::::::::::::::::::::::::::::::::::::

### Introduction

This exercise is to illustrate the concepts of selection, population size and
genetic drift, using simulations. We will use mostly 
[Red Lynx] by 
[Reed A. Cartwright](https://github.com/reedacartwright).

Another option is to
use a [web interface](https://phytools.shinyapps.io/drift-selection/) to the [R]
[learnPopGen](https://github.com/liamrevell/learnPopGen) package, but the last
one is mostly for biallelic genes (and thus not that relevant for bacteria).

Open now the [Red Lynx] website and get familiar with the different options.

You won't need the right panel (but feel free to explore). The dominance option
in the left panel won't be used either.

::::::::::::::::::::::::::::::::::::: challenge

### Genetic Drift without selection

In the first part, you will only play with the number of generations, the
initial frequency and the population size.

- Lower the number of generations to 200. 
- Adjust the population size to 1000.
- Make sure the intial frequency is set to 50%.
- Run a few simulations (ca 20).

Did any allele got fixed? What is the range of frequencies after 200 generations?

:::::::::::::::::::::::: solution

In my simulations, no allele got fixed, the final allele frequencies range 
20-80%

:::::::::::::::::::::::::::::::::

Now increase the population to 100'000, clear the graph, and repeat the 
simulations. What's the range of final frequencies now?

:::::::::::::::::::::::: solution

In my simulations, no allele got fixed, and the final allele frequencies range
45-55%

:::::::::::::::::::::::::::::::::

Now decrease the population to 10 individuals, clear the graph and repeat
these simulations. What's the range of final frequencies now?

:::::::::::::::::::::::: solution

In my simulations, one allele got fixed quickly, the latest one was at
generation 100.

:::::::::::::::::::::::::::::::::

What do you conclude here?

:::::::::::::::::::::::: solution

It is clear that stochastic (random) variation in allele frequencies 
strongly affects the probability of fixation of alleles in small populations,
not so much in large ones.

:::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: challenge

### Genetic Drift with selection

So far we've only looked at how allele frequencies vary in the absence of 
selection, that is when the two alleles provide an equal fitness. What's the
influence of random genetic drift when alleles are not neutral?

The selection strength is equivalent to the selection coefficient, i.e.
how much higher relative fitness the new allele provides. A selection 
coefficient of 0.01 means that the organism harboring the new allele has a 
1% increased fitness.

- Increase the number of generations to 1000.
- Set the selection strength to 0.01 (1%).
- Set the population size first to 100'000, then to 1000, then to 100.

How long does it take for the allele to get fixed - in average - with
the three population sizes?

:::::::::::::::::::::::: solution

About the same time, but the trajectories are much smoother with larger
populations. In the small population, it happens that the beneficial allele
disappears from the population (although not often).

:::::::::::::::::::::::::::::::::

### Fixation of slightly deleterious alleles in very small populations

We're now simulating what would happen in a very small population (or a 
population that undergoes very narrow bottlenecks), when a gene mutates. We'll
have a very small population (10 individuals), a selection value of -0.01 
(the mutated allele provides a 1% lower fitness), and a 10% initial frequency, 
which corresponds to one individual getting a mutation:

- Set the population to 10 individuals.
- Set generations to 200.
- Set initial frequency to 10%.
- Set the selection strength to -0.01.

Run many simulations. What happens?

:::::::::::::::::::::::: solution

Most new alleles go extinct quickly. In my simulations, I usually get one
of the slightly deleterious mutations fixed after 20 simulations.

:::::::::::::::::::::::::::::::::
::::::::::::::::::::::::::::::::::::::::::::::::

That's it for that part, but feel free to continue playing with the different 
settings later.

::::::::::::::::::::::::::::::::::::: keypoints

- Random genetic drift has a large influence on the probability of fixation of
alleles in small populations, even for non-neutral alleles.
- Random genetic drift has very little influence of the probability of fixation of
alleles in large populations.
- Slightly deleterious mutations can get fixed into the population through
random genetic drift, if the population is small enough and the selective value
is not too large.

::::::::::::::::::::::::::::::::::::::::::::::::

<!-- References -->

[Red Lynx]: https://cartwrig.ht/apps/redlynx/