### 2025-04-04 episodes/msa.Rmd
Corrected a div error

---
title: 'Multiple sequence alignment'
teaching: 0
exercises: 120
---

:::::::::::::::::::::::::::::::::::::: questions

- How do you align multiple sequences?
- Why is it important to properly align sequences?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Use `mafft` to align multiple sequences.
- Test different algorithms.

::::::::::::::::::::::::::::::::::::::::::::::::

## Introduction

In this episode, we are exploring multiple sequence alignment (MSA). In the first task, you are going to use [`mafft`](https://mafft.cbrc.jp/alignment/software/algorithms/algorithms.html) to align homologs of [RpoB](https://en.wikipedia.org/wiki/RpoB), the β subunit of the bacterial RNA polymerase. It is a long, multi-domain protein, suitable for showing issues related to MSA.

In the second task, you will trim that alignment to remove poorly aligned regions.

But first, go to your own folder and create a phylogenetics subfolder. You will use the alignments for other tutorials as well.

## Task 1: Use different flavors of `MAFFT` and compare the results

Start by finding the data, which is a [fasta file](episodes/files/rpoB.fasta) containing 19 sequences of RpoB, gathered by aligning RpoB from *E. coli* to a very small database present at NCBI, the [landmark](https://blast.ncbi.nlm.nih.gov/smartblast/smartBlast.cgi?CMD=Web&PAGE_TYPE=BlastDocs) database. These sequences are a subset of the sequences gathered in the previous exercise on `BLAST`.

The file is available in `/proj/g2020004/nobackup/3MK013/data/`.

:::::::::::::::::::::::::::::::::::::::  challenge

## Challenge 1.1: prepare the terrain

The course base folder is at `/proj/g2020004/nobackup/3MK013`. Go to your own folder, create a `phylogenetics` subfolder, and move into it. Also, start the `interactive` session, for 4 hours. The session name is `uppmax2025-3-4` and the cluster is `snowy`.

::: hint

Remember those commands?

```bash
ssh -Y
cd
mkdir
ls
interactive
```

::::::::

:::::::::::::::  solution

```bash
ssh -Y <username>@rackham.uppmax.uu.se
```

Remember to replace `<username>` by your name.

```bash
interactive -A <session> -M <cluster> -t <hh::mm::ss>
cd <basefolder>/<username>
mkdir <folder>
cd <folder>
```

:::::::::::::::::::::::::

:::::::::::::::  instructor

```bash
ssh -Y <username>@rackham.uppmax.uu.se
```

```bash
interactive -A uppmax2025-3-4 -M snowy -t 4:00:00
cd /proj/g2020004/nobackup/3MK013/<username>
mkdir phylogenetics
cd phylogenetics
```

:::::::::::::::::::::::::

## Challenge 1.2: copy the file

Copy the file to your newly created `phylogenetics` folder. Use a relative path.

::: hint

```bash
cp
pwd
ls ../..
```

::::::::

:::::::::::::::  solution

Use the `../../` to see what's two folders up, and then `data/rpoB/rpoB.fasta`

:::::::::::::::::::::::::

:::::::::::::::  instructor

```bash
cp ../../data/rpoB/rpoB.fasta .
```

:::::::::::::::::::::::::

<!---

```bash
cp ../../data/rpoB/rpoB.fasta .
```

--->

::::::::::::::::::::::::::::::::::::::::::::::::::

### Renaming sequences

Look at the accession ids of the fasta sequences: they are not very informative.

```bash
grep '>' rpoB.fasta | head -n 5
```

```output
>WP_000263098.1 MULTISPECIES: DNA-directed RNA polymerase subunit beta [Enterobacteriaceae]
>WP_011070610.1 DNA-directed RNA polymerase subunit beta [Shewanella oneidensis]
>WP_003114335.1 DNA-directed RNA polymerase subunit beta [Pseudomonas aeruginosa]
>WP_002228870.1 DNA-directed RNA polymerase subunit beta [Neisseria meningitidis]
>WP_003436174.1 MULTISPECIES: DNA-directed RNA polymerase subunit beta [Eubacteriales]
```

Keep the taxonomic name following the accession id, separated by `_`, using a bit of `sed` magic. Save the resulting file for later use, and show the headers again.

```bash
sed "/^>/ s/>\([^ ]*\) \([^\[]*\)\[\(.*\)\]/>\\3_\\1/"  rpoB.fasta | sed "s/ /_/g" > rpoB_renamed.fasta
grep '>' rpoB_renamed.fasta | head -n 5
```

::: discussion

## Nerd alert: dissecting the `sed` magic

This is optional reading.

The [`sed`](https://www.gnu.org/software/sed/manual/sed.html) command matches (and possibly substitutes) strings (chains of characters). In that case, the goal is to simplify the header by putting the taxonomic group first but keeping it informative enough by adding the accession number. The strategy is to match what is between the `>` and the first space, then what is between square parentheses, and put them back, separated by an underscore.

The `sed` command first matches only lines that start with a `>` (`/^>/`). It then substitutes (general pattern `s/<something>/<something else>/`) a text with another one. The first part is to match the accession id, between (escaped) square brackets, which comes after the `>` at the beginning of the line. This is expressed as `>\([^ ]*\) `: match any number of non-space characters (`[^ ]*`) and put it in memory (what is between `\(` and `\)`). Then, the description is matched by `\([^\[]*\)`, any number of characters that are not an opening bracket `[`, and put into memory. Finally, the taxonomic description is matched: `\[\(.*\)\]`, that is, any number of characters between square brackets is stored into memory. The whole line is then replaced with a `>`, the third match into memory, followed by an `_` and the content of the first match into memory `>\\3_\\1`. Then, all the input is passed through sed again, to replace any space with an underscore: `s/ /_/g` and the output is stored in a different file.

::::::::::::::

```output
>Enterobacteriaceae_WP_000263098.1
>Shewanella_oneidensis_WP_011070610.1
>Pseudomonas_aeruginosa_WP_003114335.1
>Neisseria_meningitidis_WP_002228870.1
>Eubacteriales_WP_003436174.1
```

It looks better.

### Alignment with progressive algorithm

Use `mafft` with a progressive algorithm to align the sequences.

:::::::::::::::::::::::::::::::::::::::  challenge

## Challenge 1.3: Run `mafft` with progressive algorithm

Use the FFT-NS-2 algorithm from `mafft` to align the renamed sequences. Also, record the time it takes for `mafft` to complete the task.

::: hint

Use the `module` command to load `bioinfo-tools` and `MAFFT`. Use `time` to record the time.

The help obtained through `mafft -h` is not very informative about algorithms, so check the [mafft webpage](https://mafft.cbrc.jp/alignment/software/algorithms/algorithms.html).

`mafft` actually has one executable program for each algorithm, all starting with `mafft-`. A way to display them all is to type that and push the Tab key twice to see all possibilities.

::::::::

:::::::::::::::  solution

```bash
module load <general module> <mafft module>
time mafft-<algorithm> <fasta_file> > rpoB.fftns.aln
```

```output
...
[...]
Strategy:
 FFT-NS-2 (Fast but rough)
 Progressive method (guide trees were built 2 times.)

If unsure which option to use, try 'mafft --auto input > output'.
For more information, see 'mafft --help', 'mafft --man' and the mafft page.

The default gap scoring scheme has been changed in version 7.110 (2013 Oct).
It tends to insert more gaps into gap-rich regions than previous versions.
To disable this change, add the --leavegappyregion option.

real	0m1.125s
user	0m0.818s
sys	0m0.181s
```

The last line is the output of the `time` command. It took 1.125 seconds to complete this time.

:::::::::::::::::::::::::

:::::::::::::::  instructor

```bash
module load bioinfo-tools MAFFT
time mafft-fftns rpoB_renamed.fasta > rpoB.fftns.aln
```

:::::::::::::::::::::::::

::: hint

Type `mafft` and try tab-complete to see all versions of `mafft`.

Try the command `time`

:::

::::::::::::::::::::::::::::::::::::::::::::::::::

### Alignment with iterative algorithm

Now use one of the supposedly better iterative algorithm of `mafft` to align the same sequences. Choose the [E-INS-i algorithm](https://mafft.cbrc.jp/alignment/software/algorithms/algorithms.html) which is suited for proteins that have highly conserved motifs interspersed with less conserved ones.

Take a few minutes to read upon the different alignment strategies on the page above.

:::::::::::::::::::::::::::::::::::::::  challenge

## Challenge 1.4

Use the superior E-INS-i algorithm from mafft to align the renamed sequences. Also, record the time it takes for `mafft` to complete the task.

:::::::::::::::  solution

```bash
time mafft-<better algo> <fasta file> > rpoB.einsi.aln
```

```output
[...]
Strategy:
 E-INS-i (Suitable for sequences with long unalignable regions, very slow)
 Iterative refinement method (<16) with LOCAL pairwise alignment with generalized affine gap costs (Altschul 1998)

If unsure which option to use, try 'mafft --auto input > output'.
For more information, see 'mafft --help', 'mafft --man' and the mafft page.

The default gap scoring scheme has been changed in version 7.110 (2013 Oct).
It tends to insert more gaps into gap-rich regions than previous versions.
To disable this change, add the --leavegappyregion option.

Parameters for the E-INS-i option have been changed in version 7.243 (2015 Jun).
To switch to the old parameters, use --oldgenafpair, instead of --genafpair.

real	0m7.367s
user	0m7.022s
sys	0m0.244s
```

It now took 7.36 seconds to complete this time, i.e. 6 times more than with the progressive algorithm. It doesn't make a big difference now, but with hundreds of sequences it will make one.

:::::::::::::::::::::::::

:::::::::::::::  instructor

```bash
time mafft-einsi rpoB_renamed.fasta > rpoB.einsi.aln
```

:::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::::

### Alignment visualization

You will now inspect the two resulting alignment methods. There are no convenient way to do that from Uppmax, and the easiest solution is to download the alignments **on your computer** and to use either [`seaview`](https://doua.prabi.fr/software/seaview) (you will need to install it on your computer) or an online alignment viewer like [AlignmentViewer](https://alignmentviewer.org/).

Arrange the two windows on top of each other. Change the fontsize (Props -> Fontsize in Seaview) to 8 to see a larger portion of the alignment.

:::::::::::::::::::::::::::::::::::::::  challenge

Can you spot differences? Which alignment is longer?

Hint: try to scroll to position 800-900. What do you see there? How are the blocks arranged?

:::::::::::::::  solution

Use `scp` to copy files from Uppmax to your computer. `scp` allows wildcards, but you probably need to escape the `*`.

```bash
scp <username>@rackham.uppmax.uu.se:/<absolute path to phylogenetics folder>/rpoB.\*.aln <localfolder>/
```

![](episodes/fig/seaview.png){alt='Alignments shown in seaview'}

::::::::::::::::::::::::::::::::::::::::::::::::::

:::::::::::::::  instructor

```bash
scp lionel@rackham.uppmax.uu.se:/proj/g2020004/nobackup/3MK013/<user>/phylogenetics/rpoB.\*.aln .
```

::::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::::::::::::::::

If you have time over, spend it exploring the different options of Seaview/AlignmentViewer.

## Task 2: Trim the alignment

Often, some regions of the alignment don't align properly, either because they contain low complexity segments (hinges in proteins) or evolved through repeated insertions/deletions, which alignment program cannot handle properly. It is thus good practice to remove (trim) these regions, as they are likely to worsen the quality of the subsequent phylogenetic trees. On the other hand, trimming too much of the alignment removes also potentially valuable information. There is thus a balance to be found.

In this part, you will use two different alignment trimmers, [`TrimAl`](http://trimal.cgenomics.org/) and [`ClipKIT`](https://jlsteenwyk.com/ClipKIT/), on the results of `mafft`'s E-INS-i algorithm.

### Trimming with TrimAl

First, load the module and look at the list of options available with `trimAl`. Then

```bash
module load trimAl
trimal -h
```

`trimAl` offers a lot of different options. You are going to explore two different: gap threshold (`-gt`) and the heuristic method `-automated1`, which automatically decides between three automated methods, `-gappyout`, `-strict` and `-strictplus`, based on the properties of the alignment. The gap threshold methods removes columns that contain a fraction of sequences lower than the cut-off.

For comparison purposes, you will be adding an html output.

:::::: challenge

## Challenge 2.1: TrimAl with gap threshold

Use `trimAl` to remove positions in the alignment that have more than 40% gaps.

::: solution

```bash
trimal -in rpoB.einsi.aln -out rpoB.einsi.trimalgt.aln -gt 0.6 -htmlout rpoB.einsi.trimalgt.aln.html
```

::::::::::::

::: hint

The "gap threshold" is actually expressed as fraction of non-gap residues.

::::::::

## Challenge 2.2: TrimAl with automated trimming

Use `trimAl` with the automated heuristic algorithm.

::: solution

```bash
trimal -in rpoB.einsi.aln -out rpoB.einsi.trimalauto.aln -automated1 -htmlout rpoB.einsi.trimalauto.aln.html
```

::::::::::::

## Challenge 2.3: Compare the results

Use `scp` to get the files (both alignments and html files) to your own laptop and visualize the results, by opening the `html` files with your browser and the alignment files with `seaview` or the viewer you used above.

::: solution

On your own laptop, inside the folder where you want to import the files. Replace `username` with your own username.

```bash
scp <username>@rackham.uppmax.uu.se:/proj/g2020004/nobackup/3MK013/<username>/phylogenetics/rpoB.einsi.trimal* .
```

On some OS it is necessary to escape the `*`. If the output says something about `no matches found`, try:

```bash
scp <username>@rackham.uppmax.uu.se:/proj/g2020004/nobackup/3MK013/<username>/phylogenetics/rpoB.einsi.trimal\* .
```

As before, if you do not have access to a terminal on your windows laptop, use MobaXterm and Session > SFTP to copy files to your computer.

::::::::::::

::::::::::::::::

### Trimming with ClipKIT

[`ClipKIT`](https://jlsteenwyk.com/ClipKIT/) is one of the more recent tool to trim multiple sequence alignments. In a nutshell, it tries to preseve phylogenetically-informative sites, rather than trimming gappy regions. Although it also has multiple options and modes, you will only use the default mode, `smart-gap`.

:::::: challenge

## Challenge 2.4: Use ClipKIT

To get an idea of the modes and options, look at the help of ClipKit:

```bash
module load ClipKIT
clipkit -h
```

Then run ClipKIT, explicitly using the `smart-gap` mode. Compare how much ClipKIT has trimmed the original alignment compared to trimAl.

::: solution

```bash
clipkit rpoB.einsi.aln -m smart-gap -l -o rpoB.einsi.clipkit.aln
```

```output
---------------------
| Output Statistics |
---------------------
Original length: 2043
Number of sites kept: 1543
Number of sites trimmed: 500
Percentage of alignment trimmed: 24.474%

Execution time: 0.057s
```

::::::::::::

::::::::::::::::

### Comparing all the results

Import the data and inspect the three alignments.

:::::: challenge

## Challenge 2.5: Import data and compare results

Use `scp` as above.

::: solution

```bash
scp <username>@rackham.uppmax.uu.se:/proj/g2020004/nobackup/3MK013/<username>/phylogenetics/rpoB.einsi.clipkit.aln .

```

Then use `seaview` or another viewer to visualize and compare results.

::::::::::::

::::::::::::::::

The three alignments on top of each other look like this. Click [on this link](episodes/fig/seaview_trimal_vs_clipkit.png) to better see the figure.

![](episodes/fig/seaview_trimal_vs_clipkit.png){alt='Alignments shown in seaview'}
It is of course difficult to draw conclusions based on this figure, but can you spot some trends? What alignment is more likely to generate good results?